---
title: "Exploratory Data Analysis"
author: "Michael Silva"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    toc_depth: 3
    highlight: kate
    code_folding: "hide"
---
```{r knitr_init, echo=FALSE, cache=FALSE, message=FALSE}
library(knitr)
library(rmdformats)
library(DBI)
library(dplyr)
library(ggplot2)
library(RSQLite)
library(kableExtra)

## Global options
options(max.print="150")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

## Introduction

Playbill.com publishes [the gross revenue for broadway shows by week](http://www.playbill.com/grosses).  The following data has been scrapped and extracted:

```{r}
conn <- dbConnect(RSQLite::SQLite(), "playbill.sqlite")
df <- dbGetQuery(conn, "SELECT * FROM data") %>%
  mutate(week_ending = as.Date(week_ending))
dbDisconnect(conn)
```

+ show - The name of the Broadway show
+ theatre - The theatre
+ gross - The gross revenue for the week in dollars
+ potential_gross - The potential gross revenue
+ gross_diff - The difference between the potential and actual
+ avg_ticket - Average ticket price
+ top_ticket - The maximum ticket price
+ seats_sold - The number of seats sold
+ seats_in_theatre - The total number of seats in the theatre
+ perfs - The number of performances
+ previews - The number of previews
+ percent_capacity - The percent of seats fille (seats_sold / seats_in_theatre * perfs)
+ capacity_diff - 100% - percent_capacity
+ week_ending - The date of the week ending
+ week_total_gross - The total gross revenue for all shows for the given week

## Exploration

### Summary Statistics

The following is a summary of the data.

```{r}
df %>% 
  select(-id) %>%
  summary(.) %>%
  kable() %>%
  kable_styling()
```

### Variable Distributions

The following density plots show the spread of the data.  The red verticle line is the mean and the blue verticle line is the median.

```{r}
density_plot <- function(d, title, xlab, ylab){
  ggplot(d, aes(x, stat(count))) +
    geom_density() +
    labs(title = title, 
       x = xlab,
       y = ylab) +
    theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "gray95")) +
    geom_vline(xintercept = mean(d$x), color = "red") +
    geom_vline(xintercept = median(d$x), color = "blue")
}
```

```{r}
df %>%
  mutate(x = gross / 1000000) %>%
  density_plot(., "Gross Revenue", "Millions of Dollars", "Count")
```

```{r}
df %>%
  mutate(x = potential_gross / 1000000) %>%
  density_plot(., "Potential Gross Revenue", "Millions of Dollars", "Count")
```

```{r}
df %>%
  mutate(x = gross_diff / 1000000) %>%
  density_plot(., "Gross Difference", "Millions of Dollars", "Count")
```

```{r}
df %>%
  mutate(x = avg_ticket) %>%
  density_plot(., "Average Ticket Price", "Dollars", "Count")
```

```{r}
df %>%
  mutate(x = top_ticket) %>%
  density_plot(., "Top Ticket Price", "Dollars", "Count")
```

```{r}
df %>%
  mutate(x = seats_sold) %>%
  density_plot(., "Seats Sold", element_blank(), "Count")
```

```{r}
df %>%
  mutate(x = seats_in_theatre) %>%
  density_plot(., "Seats in Theatre", element_blank(), "Count")
```

```{r}
df %>%
  mutate(x = perfs) %>%
  density_plot(., "Performances", element_blank(), "Count")
```

```{r}
df %>%
  mutate(x = previews) %>%
  density_plot(., "Previews", element_blank(), "Count")
```

```{r}
df %>%
  mutate(x = percent_capacity * 100) %>%
  density_plot(., "Capacity", "Percent", "Count")
```

### Variables Over Time

```{r}
over_time <- function(d, title, xlab, ylab){
  ggplot(d, aes(week_ending, y)) + 
  geom_point() + 
  labs(title = title, 
       x = xlab,
       y = ylab) +
  expand_limits(y=0) +
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "gray95"))
}
```

```{r}
df %>%
  mutate(y = gross / 1000000) %>%
  over_time(., "Gross Revenue by Date", element_blank(), "Millions")
```

```{r}
df %>%
  mutate(y = potential_gross / 1000000) %>%
  over_time(., "Potential Gross Revenue by Date", element_blank(), "Millions")
```

```{r}
df %>%
  mutate(y = gross_diff / 1000000) %>%
  over_time(., "Gross Revenue Difference by Date", element_blank(), "Millions")
```


```{r}
df %>%
  mutate(y = avg_ticket) %>%
  over_time(., "Average Ticket Price by Date", element_blank(), "Dollars")
```

```{r}
df %>%
  mutate(y = top_ticket) %>%
  over_time(., "Top Ticket Price by Date", element_blank(), "Dollars")
```

```{r}
df %>%
  mutate(y = seats_sold) %>%
  over_time(., "Seats Sold by Date", element_blank(), element_blank())
```

```{r}
df %>%
  mutate(y = seats_in_theatre) %>%
  over_time(., "Seats In Theatre by Date", element_blank(), element_blank())
```

```{r}
df %>%
  mutate(y = perfs) %>%
  over_time(., "Number of Preformances by Date", element_blank(), element_blank()) 
```

```{r}
df %>%
  mutate(y = previews) %>%
  over_time(., "Number of Previews by Date", element_blank(), element_blank()) 
```

```{r}
df %>%
  mutate(y = percent_capacity * 100) %>%
  over_time(., "Capacity by Date", element_blank(), element_blank()) 
```

### Aggregations

```{r}
df %>%
  group_by(week_ending) %>%
  summarise(y = n()) %>%
  over_time(., "Total Shows by Date", element_blank(), element_blank()) 
```

```{r}
df %>%
  group_by(week_ending) %>%
  summarise(y = sum(perfs)) %>%
  over_time(., "Total Preformances by Date", element_blank(), element_blank()) 
```

```{r}
df %>%
  group_by(week_ending) %>%
  summarise(y = sum(gross / 1000000)) %>%
  over_time(., "Total Gross Revenue by Date", element_blank(), "Millions")
```

```{r}
df %>%
  group_by(week_ending) %>%
  summarise(y = sum(potential_gross / 1000000)) %>%
  over_time(., "Total Potential Gross Revenue by Date", element_blank(), "Millions")
```

```{r}
df %>%
  group_by(week_ending) %>%
  summarise(y = sum(gross_diff / 1000000)) %>%
  over_time(., "Total Gross Revenue Difference by Date", element_blank(), "Millions")
```

```{r}
df %>%
  group_by(week_ending) %>%
  summarise(seats_sold = sum(seats_sold), gross = sum(gross)) %>%
  mutate(y = gross / seats_sold) %>%
  over_time(., "Overall Average Ticket Price by Date", element_blank(), "Dollars")
```

```{r}
df %>%
  group_by(week_ending) %>%
  summarize(y = mean(top_ticket)) %>%
  over_time(., "Average Top Ticket Price by Date", element_blank(), "Dollars")
```

```{r}
df %>%
  group_by(week_ending) %>%
  summarise(y = sum(seats_sold)) %>%
  over_time(., "Total Seats Sold by Date", element_blank(), element_blank())
```

```{r}
df %>%
  group_by(week_ending) %>%
  summarise(y = sum(seats_in_theatre)) %>%
  over_time(., "Total Seats In Theatre by Date", element_blank(), element_blank())
```

```{r}
df %>%
  group_by(week_ending) %>%
  summarise(y = sum(perfs)) %>%
  over_time(., "Total Preformances by Date", element_blank(), element_blank()) 
```

```{r}
df %>%
  group_by(week_ending) %>%
  summarise(y = sum(previews)) %>%
  over_time(., "Total Previews by Date", element_blank(), element_blank()) 
```

```{r}
df %>%
  group_by(week_ending) %>%
  summarise(y = mean(percent_capacity * 100)) %>%
  over_time(., "Average Capacity by Date", element_blank(), element_blank()) 
```


## Data Validation

### Complete Data Each Year

Let's check to make sure we have 52 weeks represented in each year:

```{r}
df <- df %>%
  mutate(Year = format(week_ending,"%Y"))

df %>%
  group_by(Year, week_ending) %>%
  summarise(records=n()) %>%
  mutate(count = 1) %>%
  group_by(Year) %>%
  summarise(Weeks = sum(count)) %>%
  filter(Weeks < 52) %>%
  kable() %>%
  kable_styling()
```

This checks out.  The begining and the end of the data set are incomplete but that is to be expected.  There is no data that is missing through the aquisition process.

### Closer Look at the "Zeros"

There are some zero values that need further inspection

#### Gross Revenue

There are `r nrow(df[df$gross==0,])` records that have zero gross revenue.  Let's take a closer look at these records.

```{r}
zero_gross <- df %>%
  filter(gross == 0)
```
